---
title: "How do Pokemon stats influence capture rate?"
author: "Ramray Wong"
date: "4/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
##### This dataset holds a list of Pokemon and their stats. It is different from the one used in class as it contains a more thorough description of a pokemon's stats which will be needed in this exploration. More specifically, we will be investigating how a pokemon's stats influence its capture rate. This dataset is acquired from https://www.kaggle.com/rounakbanik/pokemon titled The Complete Pokemon Dataset. It contains 801 observations -- one for each pokemon as of the making of this dataset. This dataset is tidy but later in the project, a row containing N/A was removed in order to fulfill one of the steps.   
```{r imports}
library(dplyr)
library(tidyverse)
library(ggplot2)
pokemon <- read.csv('S:/S_download/The Complete Pokemon Dataset/pokemon.csv')
pokemon <- pokemon %>%
  select(name, everything()) # move name to the beginning of the data frame

```
#### EDA
```{r basic plots}
pokemon$capture_rate <- as.numeric(pokemon$capture_rate) # convert capture rate into numeric
pokemon$is_legendary <- as.character(pokemon$is_legendary) # convert is_legendary to char
pokemon %>% # create scatter that puts capture rate vs speed.
  ggplot(aes(x = speed, y = capture_rate, color = is_legendary)) +
  geom_point()
```
*We can see that there is a negative relationship between speed and capture rate. We also notice that the legendary pokemon, regardless of speed, are largely harder to capture than other types. This is of course by design of legendary pokemon.*

```{r data frames}
pokemtx <- pokemon

# convert to numerics
pokemtx$is_legendary <- as.numeric(pokemtx$is_legendary)
pokemtx$attack <- as.numeric(pokemtx$attack)
pokemtx$defense <- as.numeric(pokemtx$defense)
pokemtx$hp <- as.numeric(pokemtx$hp)
pokemtx$speed <- as.numeric(pokemtx$speed)
pokemtx$generation <- as.numeric(pokemtx$generation)

# making correlation matrix
pokemtx <- pokemtx %>% select_if(is.numeric) %>% drop_na() # take the numerics only
# omit variables of non-interest
pokemtx2 <- pokemtx %>% select(-c(against_bug:against_water, base_egg_steps:base_total, sp_attack, sp_defense, pokedex_number, experience_growth))
```
```{r cormat}
#make a heatmap
cor(pokemtx2, use = 'pairwise.complete.obs') %>%
  as.data.frame %>% # save as data frame
  rownames_to_column %>% #convert row names to an explicit variable
  # pivot so all correlations appear on the same col
  pivot_longer(-1, names_to = 'other_var', values_to = 'correlation') %>%
  ggplot(aes(rowname, other_var, fill=correlation)) +
  # change bottom labels to be angled so they are visible
  theme(axis.text.x = element_text(angle = 40, vjust = 0.5, hjust = 1))+
  # heatmap with geom tile
  geom_tile() +
  # keep the tiles square shaped
  coord_fixed(ratio = 1) +
  # change color gradient
  scale_fill_gradient2(low='red', mid='white', high='blue')+
  #change text size and color
  geom_text((aes(label = round(correlation, 2))), color = 'black', size = 3) +
  labs(title = 'Correlation matrix for numeric variables of ', x='',y='')
```
*We can see that capture rate has correlation to be found with the other variables. For example, this matrix tells us that there is a negative correlation between capture rate and attack, defense, hp, speed, and even height and weight.*

#### MANOVA
```{r manova}
pokemanova <- manova(cbind(hp, attack, speed, defense, capture_rate) ~ type1, data = pokemon)
summary(pokemanova)
```
*We can see with significance that there were differences found among the 18 of Pokemon for at least one of the 4 different Pokemon stats (p val < 0.05, Pillai = 0.047611*

```{r anova}
summary.aov(pokemanova)
```
*Only for attack was it found that all generations differ significantly from each other (p < 0.05).*

```{r pairwise t tests}
# only attack was significant in the above part so there will be an additional t test for it
pairwise.t.test(pokemon$hp,pokemon$type1, p.adj="none")
pairwise.t.test(pokemon$attack,pokemon$type1, p.adj="none")
pairwise.t.test(pokemon$speed,pokemon$type1, p.adj="none")
pairwise.t.test(pokemon$defense,pokemon$type1, p.adj="none")
pairwise.t.test(pokemon$capture_rate,pokemon$type1, p.adj="none")
```
*From the pairwise t tests we can see that there is much significance in the difference between the mean capture rate of dragon type1's with the other type1's There is also a significant difference between mean defense stat between steel type1's and other type1's. *

*A type I error is when a true null hypothesis is rejected. In this case, it would be to reject the null hypothesis that there is no difference in the mean attack stat across the 7 generations, give this hypothesis is true.*
```{r Bonferroni}
num_tests <- 1 + 5 + 5
Prob_type_I <- 1 - (0.95)^num_tests
Bonfer <- Prob_type_I/num_tests
Prob_type_I
Bonfer
```
*The probability that a Type I error has occurred is 0.4311999. The Bonferroni significance level is 0.03919999. The post-hoc test did not become non-significant after correction.*

#### Randomization Test

```{r random sample}
# Check observed diff in defense stat between steel and normal
obs_diff <- mean(pokemon$capture_rate[pokemon$type1 == 'normal']) - 
  mean(pokemon$capture_rate[pokemon$type1 == 'dragon'])

drag_norm <- pokemon %>% filter(type1 == 'dragon' | type1 == 'normal')

set.seed(666)
mean_diff_def <- vector()
for(i in 1:5000){
  temp <- data.frame(type = drag_norm$type1, capture_rate = sample(drag_norm$capture_rate))
  mean_diff_def[i] <- temp %>%
    group_by(type) %>%
    summarise(means = mean(capture_rate)) %>%
    summarise(means_diff = diff(means)) %>%
    pull
}
```
```{r hist}
{hist(mean_diff_def, main="Distribution of the mean differences"); abline(v = -1*obs_diff, col="red")}

```
*Based on this randomization test, we can reject the null that there is no difference in mean capture rate between dragon and normal type Pokemon* 

#### Linear Regression
```{r}
pokemon_l <- pokemon

pokemon_l$attack <- pokemon_l$attack - mean(pokemon_l$attack, na.rm = TRUE)

pokemon_l$hp <- pokemon_l$hp - mean(pokemon_l$hp, na.rm = TRUE)

fit <- lm(capture_rate ~ hp + attack + attack*hp, data = pokemon_l)

summary(fit)
summary(lm(capture_rate ~ hp + attack + attack*hp, data = pokemon_l))$r.squared
```
```{r visualize plot}
ggplot(pokemon, aes(y = capture_rate,x = attack, color = hp)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  scale_color_gradient2(low='red', mid='white', high='blue') +
  xlab("Pokemon Attack stat") +
  ylab("Pokemon Capture Rate") + 
  ggtitle("Pokemon Capture Rate vs Attack stat")

```
*The mean capture rate of a Pokemon is 96.382529 if it is of 0 attack and 0 defense. Capture rate decreases by -0.863871 per unit increase in hp and decreases -0.914936 per unit increase in attack. 34.34% of the variation in the responses is explained by this model.*

```{r check assumptions}
plot(fit, which = 1)

hist(fit$residuals)

plot(fit, which = 2)


```
*We can say the normality assumption has been met because there is a decreasing pattern in the residuals, and the QQ plot sees most of the points falling on the straight line.*

```{r bootstrapping}
pokemon_nona <- na.omit(pokemon)
fit <- lm(capture_rate ~ hp + attack + hp * attack, data = pokemon_nona)

##  Bootstrap from observations
# Repeat bootstrapping 5000 times, saving the coefficients each time
samp_SEs <- replicate(5000, {
  # Bootstrap your data (resample observations)
  boot_data <- sample_frac(pokemon_nona, replace = TRUE)
  # Fit regression model
  fitboot <- lm(capture_rate ~ hp * attack, data = boot_data)
  # Save the coefficients
  coef(fitboot)
})

# Estimated SEs
samp_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd)
```
```{r}

##  Bootstrap from residuals
# Repeat bootstrapping 5000 times, saving the coefficients each time
resids_SEs <- replicate(5000, {
  # Bootstrap your residuals (resample with replacement)
  new_resids <- sample(fit$residuals, replace = TRUE)
  # Consider a new response as fitted values plus residuals
  boot_data <- pokemon_nona 
  boot_data$capture_rate = fit$fitted.values + new_resids
  # Fit regression model
  fitboot <- lm(capture_rate ~ hp * attack, data = boot_data)
  # Save the coefficients
  coef(fitboot)
})

# Estimated SEs
resids_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd)
```
*The interaction terms, hp and attack, are still significant in comparison to the original model. The standard errors for hp and attack are different from the original model but since the p value was sufficiently low, they remain significant.*

#### Logistical regression
```{r logistical}
# our binary variable is whether a pokemon is legendary. 0 means non-legendary.
pokemon$is_legendary <- as.numeric(pokemon$is_legendary)

fit_leg <- glm(is_legendary ~ capture_rate + attack, data = pokemon, family = 'binomial')
summary(fit_leg)
exp(coefficients(fit_leg))
```
*The odds that a pokemon is legendary are 0.922 times for a pokemon with a high capture rate than the odds for a pokemon with a high attack while the odds of a pokemon being legendary are 1.018 times for a pokemon with high attack than the odds are for high capture rate.*
```{r}
```
